Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-11-17 paper/main.tex
monoidal categories and functors

\begin{definition} 
  \label{def:category:F} 
  The fiber category $\mathcal{\dfiberc}$ is a product category\cite{maclaneCategoriesWorkingMathematician2013,nlab:product_category} of zero or more arbitrary categories. In the two category case, the category $\mathcal{\dfiberc}$ consists of 
  \begin{itemize}
    \item \textit{objects} ordered pairs $(x,y)$,  $x$ is an object of $\mathcal{X}$ and  $y$ is an object of $\mathcal{Y}$
    \item \textit{morphisms} ordered pairs $(f:x\rightarrow x^{\prime}, g:y\rightarrow y^{\prime})$
    \end{itemize}
The composition of morphisms is defined component wise by composition in $\mathcal{X}$ and $\mathcal{Y}$.ss
\end{definition}
The fiber category $\mathcal{\dfiber}$ can encode a large variety of collections of data fields because it is a product category of arbitrary categories. For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.  The morphisms $Hom_{\mathcal{\dfiber}}$ are functions from the data to itself, for example the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}. These functions could also be monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}.

\note{note sure this note applies anymore: We note that a one object category is a monoid \cite{nlab:monoid,barrCategoryTheoryComputing}, which is an algebraic structure that lends itself to computer library design because they are an abstraction of function composition \cite{yorgeyMonoidsThemeVariations}.}

@@

\begin{equation}
  \label{eq:artist:intersection:fiber}
  \begin{tikzcd}
    \dfiberc_{a}\times \dfiberc_{b} \arrow[d, "proj_b"] \arrow[rr, "proj_a"] &  & \dfiberc_{a} \arrow[d, "proj_c"] \\
    \dfiberc_{b} \arrow[rr, "proj_b"]                                                  &  & \dfiberc_{c}                    
    \end{tikzcd}
\end{equation}

\autoref{eq:artist:intersection:basespace} and \autoref{eq:artist:intersection:fiber} can be composed to specify how different aspects of the structure of a multivariate nested dimensional dataset, such as a spatio-temporal weather dataset, should be preserved in a visualization. 
\note{this is probably the power of this model and should maybe get a figure?}

>>>>

\begin{definition} 
  \label{def:category:F} The fiber category $\mathcal{\dfiber}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
  \begin{itemize}
    \item \textit{object} a monoid object \dfiber. A monoid is a set with a binary operator that is associative, closed, and for which the set contains an identity element $i$ \cite{nlan:monoid}.
    \item \textit{morphisms} $\dfiber \otimes \dfiber \rightarrow \dfiber$ and unit $i \rightarrow \dfiber$
  \end{itemize}
\end{definition}
The bifunctor $\otimes$ preserves identity and composition {fongInvitationAppliedCategory2019} for fibers that consist of many fields. Given a pair of objects $X, Y$, the functor yeilds a set of pairs $\{(x,y) | x \in X, y \in Y\}$. For the associated morphisms $f: X\rightarrow X$ and $g: Y\rightarrow Y$, the function $(f\times g): (X\times Y) \rightarrow (X \times Y)$ is pointwise $(f\times g)(x, y)\coloneqq (f(x), g(y))$.  This bookkeeping means that the fiber category $\mathcal{\dfiber}$ can encode a large variety of collections of data fields because it is a product category of arbitrary categories.

For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.  The morphisms $Hom_{\mathcal{\dfiber}}$ are functions from the data to itself, for example the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}. These functions could also be monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}.

@@

\begin{equation}
  \label{eq:artist:intersection:fiber}
  \begin{tikzcd}
    \dfiberc_{a}\times \dfiberc_{b} \arrow[d, "proj_b"] \arrow[rr, "proj_a"] &  & \dfiberc_{a} \arrow[d, "proj_c"] \\
    \dfiberc_{b} \arrow[rr, "proj_b"]                                                  &  & \dfiberc_{c}                    
    \end{tikzcd}
\end{equation}
%%% the fiber objects need to be composable into product categories, coproduct? -> OR on return values 
\autoref{eq:artist:intersection:basespace} and \autoref{eq:artist:intersection:fiber} can be composed to specify how different aspects of the structure of a multivariate nested dimensional dataset, such as a spatio-temporal weather dataset, should be preserved in a visualization. 
\note{this is probably the power of this model and should maybe get a figure?}
